Application of Network Flow and Linear Programming to Scheduling of Clinicians

### 2019-06-14 paper/scheduler.tex
Move docs/ -> paper/
Fix up subscript in scheduler paper
Normalize objectives of linear program
Add more details on how to normalize the objectives in writeup.

Closes #28.

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{bbm}

@@

\subsection{Constraints} \label{constraints}
	\begin{itemize}
		\item Each division needs to have exactly one clinician that covers any given block
			\begin{equation}
				\sum_{c} X_{c, b, d} = 1 \text{ for each } b, d
			\end{equation}
		\item Every weekend needs to have exactly one clinician that covers it
			\begin{equation}
				\sum_{c} Y_{c, w} = 1 \text{ for each } w
			\end{equation}
		\item In a given division, each clinician can only work between the minimum and maximum number of weeks they are allowed (note: these may be different for different divisions)
			\begin{equation}
				m_{c, d} \leq \sum_{d} X_{c, b, d} \leq M_{c, d} \text{ for each } c, d
			\end{equation}			
		\item Long weekends should be roughly equally distributed between clinicians.
			\begin{equation}
				\floor*{\frac{|\mc{L}|}{C}} \leq \sum_{w \in \mc{L}} Y_{c, w} \leq \ceil*{\frac{|\mc{L}|}{C}} \text{ for each } c
			\end{equation}
		Each clinician should work a given number of long weekends. Roughly, this is the number of long weekends divided by the number of clinicians
		\item A clinician cannot work two consecutive blocks (across all divisions)
			\begin{equation}
				\sum_d \left(X_{c, b, d} + X_{c, b+1, d}\right) \leq 1 \text{ for each $c$ and each $b \leq B - 1$}
			\end{equation}
		\item A clinician cannot work two consecutive weekends
			\begin{equation}
				Y_{c, w} + Y_{c, w+1} \leq 1 \text{ for each $c$ and each $w \leq W - 1$}
			\end{equation}
	\end{itemize}

\subsection{Objectives} \label{objectives}
Our goal in the linear program is to accommodate clinician time-off requests as much as possible. Moreover, we try to assign weekends close to blocks, if possible.
	\begin{itemize}
		\item Minimize the number of blocks assigned to a clinician that they requested as time-off
			\begin{equation}
				\label{eqn_block_obj}
				\min \sum_{c, b, d} (-1)^{ I(b \in \mc{S}_c) } \cdot X_{c, b, d}
			\end{equation}
		where $I(\cdot)$ is the indicator function.
		\item Minimize the number of weekends assigned to a clinician that they requested as time-off
			\begin{equation}
				\label{eqn_weekend_obj}
				\min \sum_{c, w} (-1)^{ I (w \in \mc{T}_c)} \cdot Y_{c, w}
			\end{equation}
		\item Maximize the adjacency between blocks and weekends assigned for a given clinician. In particular, if a clinician is assigned to work a given block in a given division, we will prefer it if the clinician also works in the weekend between the weeks of the block. \\ \\
		In our formulation, this corresponds to maximizing the sum of products of $X_{c, b, d}$ and $Y_{c, w}$ where $w$ is a weekend adjacent to block $b$ (say, $w = 2b+1$, if each block covers 2 weeks). To accomplish this, we define helper variables:
			\begin{equation}
				Z_{c, b, d} \in \{0, 1\} : \text{represents product $X_{c, b, d}\cdot Y_{c, 2b+1}$}
			\end{equation}
		with constraints
			\begin{align}
				Z_{c, b, d} &\leq X_{c, b, d} \\
				Z_{c, b, d} &\leq Y_{c, 2b+1} \text{ for each division $d$}
			\end{align}
		Thus, the objective is:
			\begin{equation}
				\label{eqn_adjacency_obj}
				\max \sum_{c, b, d} Z_{c, b, d} = \min \left(- \sum_{c, b, d} Z_{c, b, d}\right)
			\end{equation}
	\end{itemize}

We can optimize the objectives simultaneously by optimizing a weighted sum of the objective functions. We chose to weigh each objective equally.
\end{document}

>>>>

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{bbm}
\usepackage{physics}

@@

\subsection{Constraints} \label{constraints}
	\begin{itemize}
		\item Each division needs to have exactly one clinician that covers any given block
			\begin{equation}
				\sum_{c} X_{c, b, d} = 1 \text{ for each } b, d
			\end{equation}
		\item Every weekend needs to have exactly one clinician that covers it
			\begin{equation}
				\sum_{c} Y_{c, w} = 1 \text{ for each } w
			\end{equation}
		\item In a given division, each clinician can only work between the minimum and maximum number of weeks they are allowed (note: these may be different for different divisions)
			\begin{equation}
				m_{c, d} \leq \sum_{d} X_{c, b, d} \leq M_{c, d} \text{ for each } c, d
			\end{equation}
		\item Weekends should be roughly equally distributed between clinicians.
			\begin{equation}
				\floor*{ \frac{W}{C} } \leq \sum_{w} Y_{c, w} \leq \ceil*{ \frac{W}{C} } \text{ for each } c
			\end{equation}
		\item Long weekends should be roughly equally distributed between clinicians.
			\begin{equation}
				\floor*{\frac{|\mc{L}|}{C}} \leq \sum_{w \in \mc{L}} Y_{c, w} \leq \ceil*{\frac{|\mc{L}|}{C}} \text{ for each } c
			\end{equation}
		\item A clinician cannot work two consecutive blocks (across all divisions)
			\begin{equation}
				\sum_d \left(X_{c, b, d} + X_{c, b+1, d}\right) \leq 1 \text{ for each $c$ and each $b \leq B - 1$}
			\end{equation}
		\item A clinician cannot work two consecutive weekends
			\begin{equation}
				Y_{c, w} + Y_{c, w+1} \leq 1 \text{ for each $c$ and each $w \leq W - 1$}
			\end{equation}
	\end{itemize}

\subsection{Objectives} \label{objectives}
Our goal in the linear program is to accommodate clinician time-off requests as much as possible. Moreover, we try to assign weekends close to blocks, if possible.
	\begin{itemize}
		\item Minimize the number of blocks assigned to a clinician that they requested as time-off
			\begin{equation}
				\label{eqn_block_obj}
				\min_X Q_1(X) = \min_X \sum_{c, b, d} (-1)^{ I(b \in \mc{S}_c) } \cdot X_{c, b, d}
			\end{equation}
		where $I(\cdot)$ is the indicator function.
		\item Minimize the number of weekends assigned to a clinician that they requested as time-off
			\begin{equation}
				\label{eqn_weekend_obj}
				\min_Y Q_2(Y) = \min_Y \sum_{c, w} (-1)^{ I (w \in \mc{T}_c)} \cdot Y_{c, w}
			\end{equation}
		\item Maximize the adjacency between blocks and weekends assigned for a given clinician. In particular, if a clinician is assigned to work a given block in a given division, we will prefer it if the clinician also works in the weekend between the weeks of the block. \\ \\
		In our formulation, this corresponds to maximizing the sum of products of $X_{c, b, d}$ and $Y_{c, w}$ where $w$ is a weekend adjacent to block $b$ (say, $w = 2b+1$, if each block covers 2 weeks). To accomplish this, we define helper variables:
			\begin{equation}
				Z_{c, b, d} \in \{0, 1\} : \text{represents product $X_{c, b, d}\cdot Y_{c, 2b+1}$}
			\end{equation}
		with constraints
			\begin{align}
				Z_{c, b, d} &\leq X_{c, b, d} \\
				Z_{c, b, d} &\leq Y_{c, 2b+1} \text{ for each division $d$}
			\end{align}
		Thus, the objective is:
			\begin{equation}
				\label{eqn_adjacency_obj}
				\max_Z \sum_{c, b, d} Z_{c, b, d} = \min_Z \left(- \sum_{c, b, d} Z_{c, b, d}\right) = \min_Z Q_3(Z)
			\end{equation}
	\end{itemize}

In order to optimize all the objectives simultaneously we can use the weighted sum method which guarantees our solutions to be Pareto optimal (??). It is necessary to normalize the objective functions, since their domains are not equivalent. In general, if the objective function is of the from
\[
 	f(\vb{x}) = \sum_i a_ix_i
\]
Then we normalize it by dividing by the sum of absolute weights, thus giving
\[
	\bar{f}(\vb{x}) = \frac{f(\vb{x})}{\sum_i \abs{a_i}}
\]
In our case, this gives us the following objective functions:
	\begin{align*}
		\bar{Q}_1(X) = \frac{Q_1(X)}{\sum_{c, b, d} 1} = \frac{Q_1(X)}{C \cdot B \cdot D} \\
		\bar{Q}_2(Y) = \frac{Q_2(Y)}{\sum_{c, w} 1} = \frac{Q_2(Y)}{C \cdot W} \\
		\bar{Q}_3(Z) = \frac{Q_1(Z)}{\sum_{c, b, d} 1} = \frac{Q_3(Z)}{C \cdot B \cdot D}
	\end{align*}
We also chose to weigh the objectives equally, thus giving us the final objective:
	\begin{equation}
		\min_{X, Y, Z} \frac{\bar{Q}_1(X) + \bar{Q}_2(Y) + \bar{Q}_3(Z)}{3}
	\end{equation}
\end{document}
